---
title: "Spatial Join of Asset Data for Climate Change Vulnerability Assessment"
format: html
author: "Jordan Sibley"
date: "7/20/2026"
---

## Set up

```{r}
library(tidyverse)
library(sf)
library(readxl)
library(leaflet)
library(writexl)
library(testthat)

library(RecordLinkage) # for the string matching 
```

```{r}
# Data File paths

gdb_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/CCVA-asset-classifcation.gdb"

fc_name <- "GOGA_FACILITY_BuildingsAndStructures_py_20260520_ExportFeatures"

excel_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned.xlsx"

output_excel <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned-with-flag.xlsx"
```

## Read in data

### Asset polygon layer with cultural resource information

```{r}
# Read polygon layer
buildings <- st_read(
  dsn = gdb_path,
  layer = fc_name,
  quiet = FALSE
)

# Transform CRS to WGS84 
buildings <- buildings |> 
  st_transform(4326) 

# Check field names
names(buildings)
```

```{r}
# REMOVE DECOMISSIONED BUILDINGS ----
# Clean building data to match definition query in layer package (only decomissioned o start, the other part will be filtered before the spatial join )
buildings <- buildings |> 
  # Filter to buildings in park
  filter(InPark == "Yes") |> 
  # filter out Decommissioned
  filter(!BLDGSTATUS %in% c("Decommissioned")) 

# PREP FOR JARO-WINKLER SIMILARITY TEST ----
# Created cleaned name column in buildings 
buildings <- buildings |> 
  # duplicate the RESNAME column 
  mutate(RESNAME_CLEAN = RESNAME) |> 
   # convert text to all upper case
  mutate(RESNAME_CLEAN = toupper(RESNAME_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(RESNAME_CLEAN = str_replace_all(RESNAME_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(RESNAME_CLEAN = str_squish(RESNAME_CLEAN)) |> 
  # relocate column for comparison
  relocate(RESNAME_CLEAN, .after = RESNAME)
```

### GOGA asset tabular data

```{r}
# Read asset spreadsheet
assets <- read_excel(excel_path, sheet = "Sheet1")

# Remove characters (+) from lat and lon columns and convert to numeric
assets <- assets |> 
  mutate(
    PrimaryLatitude = gsub("\\+", "", PrimaryLatitude),
    PrimaryLongitude = gsub("\\+", "", PrimaryLongitude))|>
  mutate(PrimaryLatitude = as.numeric(PrimaryLatitude),
         PrimaryLongitude = as.numeric(PrimaryLongitude)) |> 
  # rename Location to LocationID
  rename(LocationID = Location)

# Create rowID column to maintain row order at the end
assets <- assets |> 
  mutate(RowID = row_number())


# PREP FOR JARO-WINKLER SIMILARITY TEST ----
# Created cleaned name column in assets
assets <- assets |> 
  # duplicate the Description column
  mutate(Description_CLEAN = Description) |> 
  # convert text to all upper case
  mutate(Description_CLEAN = toupper(Description_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(Description_CLEAN = str_replace_all(Description_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(Description_CLEAN = str_squish(Description_CLEAN)) |> 
  # relocate column for comparison
  relocate(Description_CLEAN, .after = Description)
```

```{r}
# For visual check 
# Filter out rows that are missing lat lon points 
assets_point <- assets |> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))

# Convert spreadsheet coordinates to sf points
assets_sf <- st_as_sf(
  assets_point,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_sf <- st_transform(
  assets_sf,
  st_crs(buildings)
)
```

## Visual Check

```{r}
# Get buildings dataset in form for leaflet map
buildings_leaflet <- buildings |>
  st_zm() |>
  st_cast("MULTIPOLYGON") 

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_sf,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```

## Join by FACLOCID number

```{r}
# Make polygons valid 
buildings_fixed <- st_make_valid(buildings_leaflet)

# Clean FACLOCID in buildings_fixed (to account for the duplicate id numbers that appear)
# From an examination and cross checking the first number in the series is the current ID
buildings_fixed <- buildings_fixed |> 
  mutate(FACLOCID = sub(";.*", "", FACLOCID))

# Rename LocationID column to be FACLOCID for join
assets_facloc <- assets |>
  rename(FACLOCID = LocationID)

# Perform left join to match asset data to CONTRIBRES column by matching ID 
assets_facloc_joined <- assets_facloc |>
  left_join(buildings_fixed |>
              st_drop_geometry() |>
              # include RESNAME_CLEAN and BLDGALTNAME for JW check
              select(FACLOCID, CONTRIBRES, RESNAME_CLEAN, BLDGALTNAME), 
            by = "FACLOCID")

# View how many tagged
sum(!is.na(assets_facloc_joined$CONTRIBRES))
```

### Check FACLOCID join using a Jaro-Winkler Similarity test

In order to better confirm that the LocationID/FACLOCID match actually matched the correct assets together, the names of the buildings in the two datasets (Description_CLEAN and RESNAME_CLEAN) will be compared for a given row. The '\_CLEAN' version of the building names will be used as it works better with the JW test

```{r}
# Filter assets_facloc_joined to rows where CONTRIBRES is not NA
assets_facloc_check <- assets_facloc_joined |>
  filter(!is.na(CONTRIBRES))

# Calculate Jaro-Winkler similarity between Description_CLEAN and RESNAME_CLEAN
assets_facloc_check <- assets_facloc_check |>
  mutate(jw_score = mapply(RecordLinkage::jarowinkler, Description_CLEAN, RESNAME_CLEAN))

# Filter to the Jaro-Winkler score of less than 0.90 for review
assets_facloc_review <- assets_facloc_check |>
  filter(jw_score < 0.90) |>
  select(
    FACLOCID,
    Description_CLEAN,
    RESNAME_CLEAN,
    jw_score,
    everything()
  )

View(assets_facloc_review)
```

FACLOCID numbers to flag: 40834 Description: PARKING LOT BATTERY ALEXANDER Resname: QUARTERS 804A

Found in data of this asset: "This feature is incorrectly described as "Battery Alexander parking lot" in FMSS."

```{r}
# Remove CONTRIBRES label for FACLOCID = 40834
assets_facloc_joined <- assets_facloc_joined |> 
  mutate(CONTRIBRES = case_when(
    FACLOCID == "40834" ~ NA_character_, 
    TRUE ~ CONTRIBRES
  ))
```

## Spatial join for the untagged set

Note for self: This join will remove rows that do not contain lat lon points. I should be okay when I recombine everything at the end because these rows are included in the first ID join.

```{r}
# Clean buildings dataset to not include planned or proposed buildings 
buildings_fixed_current <- buildings_fixed |> 
  # Planned and proposed
   filter(!BLDGSTATUS %in% c("Proposed","Planned")) 
```

```{r}
# Filter to rows that do not contain contribres tag
assets_needing_spatial <- assets_facloc_joined |>
  filter(is.na(CONTRIBRES)) |> 
  # remove CONTRIBRES, RESNAME_CLEAN, BLDGALTNAME
  select(c(-CONTRIBRES, -RESNAME_CLEAN, -BLDGALTNAME))


# Turn assets data into spatial points ----
# Filter out rows that are missing lat lon points 
assets_needing_spatial <- assets_needing_spatial|> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))

# Convert spreadsheet coordinates to sf points
assets_spatial_sf <- st_as_sf(
  assets_needing_spatial,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_spatial_sf <- st_transform(
  assets_spatial_sf,
  st_crs(buildings_fixed)
)

# Perform spatial join 
assets_spatial_join <- st_join(
  assets_spatial_sf,
  buildings_fixed_current |> 
    # include CONTRIBRES, RESNAME_CLEAN, BLDGALTNAME
    select(CONTRIBRES, RESNAME_CLEAN, BLDGALTNAME),
  join = st_within,
  left = TRUE
)

# View how many are still untagged (there should def be a decent amount from visual check in maps)
sum(!is.na(assets_spatial_join$CONTRIBRES))

```

### Check Spatial using a Jaro-Winkler Similarity test

To check if the spatial join of points within building polygons accurately matched the buildings, the names of the buildings in the two datasets (Description_CLEAN and RESNAME_CLEAN) will be compared for a given row that received a tag. The '\_CLEAN' version of the building names will be used as it works better with the JW test

```{r}
# Filter assets_facloc_joined to rows where CONTRIBRES is not NA
assets_spatial_check <- assets_spatial_join |>
  filter(!is.na(CONTRIBRES))

# Calculate Jaro-Winkler similarity between Description_CLEAN and RESNAME_CLEAN
assets_spatial_check <- assets_spatial_check |>
  mutate(jw_score = mapply(RecordLinkage::jarowinkler, Description_CLEAN, RESNAME_CLEAN))

# Filter to the Jaro-Winkler score of less than 0.90 for review
assets_spatial_review <- assets_spatial_check |>
  filter(jw_score < 0.90) |>
  select(
    jw_score,
    Description_CLEAN,
    RESNAME_CLEAN,
    BLDGALTNAME,
    CONTRIBRES,
    FACLOCID,
    everything()
  )

View(assets_spatial_review)
```

Rows to flag (change CONTRIBRES to NA)

38390 "Restroom, Roadway (AL-)" tagged: Paved area

38371 "Guardhouse AL-22" tagged: Roadway (wharf to NW)

240697 "Photovoltaic system; Roof, Pier 2" tagged: Pier 2

80793 "Slide Ranch Farm to Table Teaching Center (SR-1)" tagged: Bath house

81751 "Fire Supression System, Alcatraz AL-" tagged: Alcatrax cellhouse

80093 "Parking lot, Marin Headlands VC (RN-P950)" tagged: VISITOR CENTER FORMER

111239 "MUWO Phone System" tagged: ADMINISTRATION CONCESSION BUILDING

38269 "Electrical Distribution Sytem, Fort Mason" tagged: community garden

226326 "Photovoltaic system; Roof, FM-201" tagged: Fort Mason Headquarters

81745 "Electrical Distribution System, Alcatraz (AL-)" tagged: Roadway (wharf to NW)

```{r}
# Change CONTRIBRES label for flagged rows ----

# Create look up table of flagged rows 
rows_to_flag <- tribble(
  ~FACLOCID, ~Description,
  "38390",  "Restroom, Roadway (AL-)",
  "38371",  "Guardhouse AL-22",
  "38269",  "Electrical Distribution Sytem, Fort Mason",
  "111239", "MUWO Phone System",
  "240697", "Photovoltaic system; Roof, Pier 2",
  "80793",  "Slide Ranch Farm to Table Teaching Center (SR-1)",
  "81751",  "Fire Supression System, Alcatraz AL-",
  "80093",  "Parking lot, Marin Headlands VC (RN-P950)",
  "226326", "Photovoltaic system; Roof,  FM-201",
  "81745",  "Electrical Distribution System, Alcatraz (AL-)"
)

# Clean spatial join dataset
assets_spatial_join <- assets_spatial_join |> 
  left_join(rows_to_flag |>  mutate(flag_na = TRUE),
            by = c("FACLOCID", "Description")) |> 
  mutate(CONTRIBRES = if_else(flag_na, NA_character_, CONTRIBRES)) |> 
  select(-flag_na)

```


## Recombine and clean data

Methods: The dataset created after the first join 'assets_facloc_joined' should contain all the orginial rows (nothing filtered out). It does however, contain more rows than the orginal dataset due to 'buildings_fixed' containing multiple rows with the same FACLOCID (from a code check it appears that there are 40 FACLOCID numbers that show up more than once). Additionally, the spatial join resulted in a few duplicated rows due to an overlaping of polygons. To solve for this problem, once the data is recombined, I will condense the data by the orginal rowID and will see if any rows now contain different information for CONTRIBRES. 

```{r}
# Drop geometry from spatial join 
assets_spatial_join <- assets_spatial_join |> 
  st_drop_geometry()

# Bind rows of assets_spatial_join and assets_facloc_joined
full_join_unclean <- bind_rows(assets_facloc_joined, assets_spatial_join)

# Create a summary of responses for CONTRIBRES
duplicate_summary <- full_join_unclean |> 
  group_by(RowID) |> 
  summarise(CONTRIBRES_options = paste(unique(CONTRIBRES),
                                       collapse = "; "),
            n_options = n_distinct(CONTRIBRES),
            .groups = "drop")

# Create a single record per RowID number 
full_join_correct_row <- full_join_unclean |> 
  group_by(RowID) |> 
  summarise(across(everything(), first),
            .groups = "drop")

# Add the list of possible values for contribres
full_join_correct_row <- full_join_correct_row |> 
  left_join(duplicate_summary, 
            by = "RowID")

# Create ambiguous cases column 
full_join_correct_row <- full_join_correct_row |>
  mutate(CONTRIBRES = if_else(n_options > 1,
                              "Multiple responses",
                              CONTRIBRES),
         CONTRIBRES_options = if_else(n_options > 1,
                                      CONTRIBRES_options,
                                      NA_character_))
```

The resulting dataset only shows 2 rows that contain multiple responses for CONTRIBRES, both of which are Unknown;No. I will change both of these rows to be 'No'

```{r}
# Change ambiguous responses to No
full_join_correct_row <- full_join_correct_row |> 
  mutate(CONTRIBRES = if_else(CONTRIBRES == "Multiple responses",
                              "No",
                              CONTRIBRES))

# Clean dataset by removing and renaming columns 
full_join_clean <- full_join_correct_row |> 
  # remove columns from building dataset
  select(c(-Description_CLEAN, -RESNAME_CLEAN, -BLDGALTNAME, -CONTRIBRES_options, -n_options)) |> 
  # rename columns 
  rename("LocationID" = "FACLOCID", 
         "ContributingResourceFlag" = "CONTRIBRES") |> 
  # confirm correct order 
  arrange(RowID) |> 
  # remove RowID
  select(c(-RowID))
```


## QAQC and export 
```{r}
# Check that the row number in orginal assets dataset matches final dataset
# Recheck row count
 test_that("Final dataset has same number of rows as original asset dataset", {
   expect_equal(nrow(full_join_clean), nrow(assets))
 })

# Write final dataset to Excel
write_xlsx(full_join_clean, output_excel)

cat("Done. Output saved to:", output_excel)

```

