---
title: "Spatial Join of Asset Data for Climate Change Vulnerability Assessment"
format: html
author: "Jordan Sibley"
date: "7/20/2026"
---

## Set up

```{r}
library(tidyverse)
library(sf)
library(readxl)
library(leaflet)
library(writexl)
library(testthat)

library(RecordLinkage) # for the string matching 
```

```{r}
# Data File paths

gdb_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/CCVA-asset-classifcation.gdb"

fc_name <- "GOGA_FACILITY_BuildingsAndStructures_py_20260520_ExportFeatures"

excel_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned.xlsx"

output_excel <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned-with-flag.xlsx"
```

## Read in data

### Asset polygon layer with cultural resource information

```{r}
# Read polygon layer
buildings <- st_read(
  dsn = gdb_path,
  layer = fc_name,
  quiet = FALSE
)

# Transform CRS to WGS84 
buildings <- buildings |> 
  st_transform(4326) 

# Check field names
names(buildings)
```

```{r}
# Clean building data to match definition query in layer package
buildings <- buildings |> 
  # Filter to buildings in park
  filter(InPark == "Yes") |> 
  # filter out Decommissioned
  filter(!BLDGSTATUS %in% c("Decommissioned")) #|> 
  # Planned and proposed
 #  filter(!BLDGSTATUS %in% c("Proposed","Planned")) 
```

### GOGA asset tabular data

```{r}
# Read asset spreadsheet
assets <- read_excel(excel_path, sheet = "Sheet1")

# Remove characters (+) from lat and lon columns and convert to numeric
assets <- assets |> 
  mutate(
    PrimaryLatitude = gsub("\\+", "", PrimaryLatitude),
    PrimaryLongitude = gsub("\\+", "", PrimaryLongitude))|>
  mutate(PrimaryLatitude = as.numeric(PrimaryLatitude),
         PrimaryLongitude = as.numeric(PrimaryLongitude)) |> 
  # rename Location to LocationID
  rename(LocationID = Location)

# Create rowID column to maintain row order at the end
assets <- assets |> 
  mutate(RowID = row_number())

# Filter out rows that are missing lat lon points 
assets_point <- assets |> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))
```

```{r}
# Convert spreadsheet coordinates to sf points
assets_sf <- st_as_sf(
  assets_point,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_sf <- st_transform(
  assets_sf,
  st_crs(buildings)
)
```

### Finding ways to fill in missing data

```{r}
# MIGHT NOT USE
goga_buildings_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/GOGA_Buildings.gdb"

GOGA_buildings <- st_read(
  dsn = goga_buildings_path,
  layer = "GOGA_FACILITY_Buildings_py",
  quiet = FALSE
)

# Check field names
names(GOGA_buildings)
```

```{r}
# Identify rows with missing lat lon points 
assets_missing_point <- assets |> 
  filter(is.na(PrimaryLatitude) | is.na(PrimaryLongitude))

# See if missing rows are in GOGA_Building via Location ID
matches_GOGA_build <- assets_missing_point |> 
  filter(LocationID %in% GOGA_buildings$FACLOCID)

# See if missing rows are in buildings dataset via Location ID
matches_buildings <- assets_missing_point |> 
  filter(LocationID %in% buildings$FACLOCID)
```

80 rows within assets dataset do not have a lat lon point. Of those 80, only 7 have a matching Location ID within GOGA_Buildings dataset and 1 has a matching Location ID within buildings dataset. Need to decide if its worth it? Might be wiser to go through and assign lat lon point to missing rows

## Visual Check

```{r}
# Get buildings dataset in form for leaflet map
buildings_leaflet <- buildings |>
  st_zm() |>
  st_cast("MULTIPOLYGON") 

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_sf,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```

## Join by FACLOCID number

```{r}
# Make polygons valid 
buildings_fixed <- st_make_valid(buildings_leaflet)

# Clean FACLOCID in buildings_fixed (to account for the duplicate id numbers that appear)
# From an examaninaton and cross checking the first number in the series is the current ID
buildings_fixed <- buildings_fixed |> 
  mutate(FACLOCID = sub(";.*", "", FACLOCID))

# Rename LocationID column to be FACLOCID for join
assets_facloc <- assets_sf |>
  rename(FACLOCID = LocationID)

# Perform left join to match asset data to CONTRIBRES column by matching ID 
assets_facloc_joined <- assets_facloc |>
  left_join(buildings_fixed |>
              st_drop_geometry() |>
              select(FACLOCID, CONTRIBRES),
            by = "FACLOCID")

# View how many tagged
sum(!is.na(assets_facloc_joined$CONTRIBRES))
```

424 asset rows were matched with contributing resource tag from matching FACLOCID number

NOTE: Only 399 assets were matched once buildings dataset took on definition query. From a check in assets dataset, 52 assets are of Status = PLANNED. In a check of contributing resource flag within buildings, all planned or proposed buildings did not have a contributing resource flag == yes (most were no). I think because of that reason, I will wait to remove these polygons until after this join to capture these likely No labels.

When I re ran the code when only filtered out Decommissioned assets in buildings dataset, it matched 402 assets by FACLOCID.

```{r}
# Recreate map with only missed tagged points 

# Filter to rows that do not contain contribres tag
assets_needing_spatial <- assets_facloc_joined |>
  filter(is.na(CONTRIBRES))

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_needing_spatial,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```

There are still plenty of points that could be matched by the spatial join.

## Testing out a join via approximate string matches for asset names

First attempt: Jaro-winkler

Jaro-Winkler Similarity measures the number of characters between strings that need to be edited for the strings to match, while giving more weight to the first half of the strings. I think this strategy could be helpful for the weird set of numbers that appears in some of the asset names in assets_sf (ex.(PE1630))

My thought is that we could first use it to find exact names, then see when the score is higher than 0.8, and then also maybe use it as a way to flag items within the spatial join?

Data: For this trial I am going to use assets$Descriptions as one group and buildings_fixed$RESNAME as the other. Which means this is all of the all fo the assets, even the ones that do not contain lat lon points in hope that we can find a match

```{r}
# Write function to calculate Jaro-Winkler similailty using RecordLinkage package
approxMatch = function(base_series, reference_series){
  # base_series = column of names in one dataset
  # reference_series = column of names in other dataset
  # function returns a dataframe in which: 
  # 1st column is base_series text 
  # 2nd column contains what the algorithm believes is the match based on J-W similarity score
  # 3rd column is J-W score
  # 4th column indicates whether the match is an exact(TRUE) or approximate (FALSE) match
  base_series = sort(unique(base_series)) # filter only unique items from one of the series 
  reference_series = sort(unique(reference_series)) # and sort alphabetically
  
  matches = as.data.frame(base_series)
  matches$matched = NA # create empty column of matched base_series
  matches$similarity = NA
  for (i in 1:nrow(matches)){
    sim_scores = list()
    for (building_name in reference_series){
      sim_scores = append(sim_scores, jarowinkler(matches$base_series[i], building_name))
    } # loop through base series and create list of Jaro-Winkler similarity scores 
    matches$similarity[i] = max(unlist(sim_scores)) # return the highest similarity scores
    matches$matched[i] = reference_series[which(sim_scores == max(unlist(sim_scores)))]
    # take the item name corresponding to the highest similarity score 
  }
  matches$absolute = ifelse(matches[,1] == matches$matched, TRUE, FALSE)
  return(matches)
}

```

```{r}
# Test function 
matches = approxMatch(assets$Description, buildings_fixed$RESNAME)
View(matches)
```

First impressions are that this might work! Confirmed that it gets the exact matches (score = 1.0). Might need to look into why when a match is not made (reference_series = blank), it still says that the similarity score is ~ 0.6722222

Second attempt: Clean names before running function
- lower case all words
- remove punctuation
- XX-XXX
- RNXXX
FBR-XXX

```{r}
# Created cleaned name column in assets
assets <- assets |> 
  # duplicate the Description column
  mutate(Description_CLEAN = Description) |> 
  # convert text to all upper case
  mutate(Description_CLEAN = toupper(Description_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(Description_CLEAN = str_replace_all(Description_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(Description_CLEAN = str_squish(Description_CLEAN)) |> 
  # relocate column for comparison
  relocate(Description_CLEAN, .after = Description)

# Created cleaned name column in buildings 
buildings_fixed <- buildings_fixed |> 
  # duplicate the RESNAME column 
  mutate(RESNAME_CLEAN = RESNAME) |> 
   # convert text to all upper case
  mutate(RESNAME_CLEAN = toupper(RESNAME_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(RESNAME_CLEAN = str_replace_all(RESNAME_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(RESNAME_CLEAN = str_squish(RESNAME_CLEAN)) |> 
  # relocate column for comparison
  relocate(RESNAME_CLEAN, .after = RESNAME)
```

```{r}
# Test function after cleaning 
matches_post_cleaning = approxMatch(assets$Description_CLEAN, buildings_fixed$RESNAME_CLEAN)
View(matches_post_cleaning)
```

This version of running the function was a lot more successful as each asset in assets dataset was matched with a name in buildings. 

## Spatial join for the untagged set

```{r}
# Perform spatial join 
assets_spatial <- st_join(
  assets_needing_spatial,
  buildings_fixed |> 
    select(CONTRIBRES),
  join = st_within,
  left = TRUE
)

# Combine 2 CONTRIBRES columns 
assets_spatial <- assets_spatial |>
  mutate(CONTRIBRES = coalesce(CONTRIBRES.x, CONTRIBRES.y)) |> 
  # remove x and y column 
  select(c(-CONTRIBRES.x, -CONTRIBRES.y))

# View how many are still untagged (there should def be a decent amount from visual check in maps)
sum(!is.na(assets_spatial$CONTRIBRES))

```

Only 332 rows out of 1387 (23%) are missing the contributing resource tag.

## Recombine data

```{r}
# Get rows from left join that did not result in NA value
assets_facloc_tagged <- assets_facloc_joined |>
  filter(!is.na(CONTRIBRES))

# Join with spatial join (which also contains the rows that did not receive a tag)
assets_sp_left_join <- bind_rows(assets_facloc_tagged, assets_spatial)

# Rename FACLOCID to LocationID and drop geometry
assets_sp_left_join <- assets_sp_left_join |> 
  rename(LocationID = FACLOCID) |> 
  st_drop_geometry()

# Add back the rows that did not include lat lon points and reorder by RowID
assets_final <- bind_rows(assets_sp_left_join, assets_missing_point) |>
  arrange(RowID)

# Confirm final asset data contains the same number of rows are original asset data
# test_that("Final dataset has same number of rows as original asset dataset", {
#   expect_equal(nrow(assets_final), nrow(assets))
# })
# COMMENTED OUT FOR RUNNING PURPOSES
```

Error message: ── Failure: Final dataset has same number of rows as original asset dataset ────────────────────────────────────────────────────────────── Expected `nrow(assets_final)` to equal `nrow(assets)`. Differences: 1/1 mismatches \[1\] 1442 - 1387 == 55

## Figure out error

```{r}
# Find duplicate FACLOCID 
buildings_fixed |>
  st_drop_geometry() |>
  count(FACLOCID) |>
  filter(n > 1) |>
  arrange(desc(n))

```

There are 44 different FACLOCID numbers that appear more than once

```{r}
# Did the FACLOCID join increase the row count?
nrow(assets_facloc)
nrow(assets_facloc_joined)

# Did the spatial join increase the row count?
nrow(assets_needing_spatial)
nrow(assets_spatial)
```

The FACLOCID join increased the row count by 39 The spatial join increased the row count by 16

## Fix duplicated rows from the joins

```{r}
# Find rows that got duplicated
dup_rowids <- assets_final |>
  count(RowID) |>
  filter(n > 1)

# Inspect what is happening
assets_final |>
  semi_join(dup_rowids, by = "RowID") |>
  arrange(RowID) |> 
  select(RowID, Description, CONTRIBRES)
```

It seems some of them were duplicated due to multiple values for CONTRIBRES, it also seems that things got messed up due to some buildings which contain separate polygons contain the same FACLOCID and from the visual check there are some overlaping polygons

```{r}
# Create a summary for all responses
duplicate_summary <- assets_final |>
  group_by(RowID) |>
  summarise(
    CONTRIBRES_options = paste(
      unique(CONTRIBRES),
      collapse = "; "
    ),
    n_options = n_distinct(CONTRIBRES),
    .groups = "drop"
  )

# Create a single record per RowID
assets_final_dedup <- assets_final |>
  group_by(RowID) |>
  summarise(
    across(everything(), first),
    .groups = "drop"
  )

# Add the list of possible values
assets_final_dedup <- assets_final_dedup |>
  left_join(
    duplicate_summary,
    by = "RowID"
  )

# Flag ambiguous cases column 
assets_final_dedup <- assets_final_dedup |>
  mutate(CONTRIBRES = if_else( n_options > 1,
                               "Multiple responses",
                               CONTRIBRES),
         CONTRIBRES_options = if_else( n_options > 1,
                                       CONTRIBRES_options,
                                       NA_character_)
         )

# Recheck row count
test_that("Final dataset has same number of rows as original asset dataset", {
  expect_equal(nrow(assets_final_dedup), nrow(assets))
})
```

## Final clean and export

```{r}
# Clean columns for export
assets_final_dedup <- assets_final_dedup |> 
  # remove n_options and RowID column
  select(c(-n_options, -RowID)) |> 
  # rename CONTRIBRES to ContributingResourceFlag
  rename(ContributingResourceFlag = CONTRIBRES,
         FlagOptions = CONTRIBRES_options)
```

```{r}
# Write result to Excel
# write_xlsx(
#   assets_final_dedup,
#   output_excel
# )
# 
# cat("Done. Output saved to:", output_excel)
```



# FRESH START -----

## Set up

```{r}
library(tidyverse)
library(sf)
library(readxl)
library(leaflet)
library(writexl)
library(testthat)

library(RecordLinkage) # for the string matching 
```

```{r}
# Data File paths

gdb_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/CCVA-asset-classifcation.gdb"

fc_name <- "GOGA_FACILITY_BuildingsAndStructures_py_20260520_ExportFeatures"

excel_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned.xlsx"

output_excel <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned-with-flag.xlsx"
```

## Read in data

### Asset polygon layer with cultural resource information

```{r}
# Read polygon layer
buildings <- st_read(
  dsn = gdb_path,
  layer = fc_name,
  quiet = FALSE
)

# Transform CRS to WGS84 
buildings <- buildings |> 
  st_transform(4326) 

# Check field names
names(buildings)
```

```{r}
# REMOVE DECOMISSIONED BUILDINGS ----
# Clean building data to match definition query in layer package (only decomissioned o start, the other part will be filtered before the spatial join )
buildings <- buildings |> 
  # Filter to buildings in park
  filter(InPark == "Yes") |> 
  # filter out Decommissioned
  filter(!BLDGSTATUS %in% c("Decommissioned")) 

# PREP FOR JARO-WINKLER SIMILARITY TEST ----
# Created cleaned name column in buildings 
buildings <- buildings |> 
  # duplicate the RESNAME column 
  mutate(RESNAME_CLEAN = RESNAME) |> 
   # convert text to all upper case
  mutate(RESNAME_CLEAN = toupper(RESNAME_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(RESNAME_CLEAN = str_replace_all(RESNAME_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(RESNAME_CLEAN = str_squish(RESNAME_CLEAN)) |> 
  # relocate column for comparison
  relocate(RESNAME_CLEAN, .after = RESNAME)
```

### GOGA asset tabular data

```{r}
# Read asset spreadsheet
assets <- read_excel(excel_path, sheet = "Sheet1")

# Remove characters (+) from lat and lon columns and convert to numeric
assets <- assets |> 
  mutate(
    PrimaryLatitude = gsub("\\+", "", PrimaryLatitude),
    PrimaryLongitude = gsub("\\+", "", PrimaryLongitude))|>
  mutate(PrimaryLatitude = as.numeric(PrimaryLatitude),
         PrimaryLongitude = as.numeric(PrimaryLongitude)) |> 
  # rename Location to LocationID
  rename(LocationID = Location)

# Create rowID column to maintain row order at the end
assets <- assets |> 
  mutate(RowID = row_number())


# PREP FOR JARO-WINKLER SIMILARITY TEST ----
# Created cleaned name column in assets
assets <- assets |> 
  # duplicate the Description column
  mutate(Description_CLEAN = Description) |> 
  # convert text to all upper case
  mutate(Description_CLEAN = toupper(Description_CLEAN)) |> 
  # remove punctuation and replace with an empty space
  mutate(Description_CLEAN = str_replace_all(Description_CLEAN,"[^A-Z0-9 ]", " ")) |> 
  # remove leading and extra white space
  mutate(Description_CLEAN = str_squish(Description_CLEAN)) |> 
  # relocate column for comparison
  relocate(Description_CLEAN, .after = Description)
```


```{r}
# For visual check 
# Filter out rows that are missing lat lon points 
assets_point <- assets |> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))

# Convert spreadsheet coordinates to sf points
assets_sf <- st_as_sf(
  assets_point,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_sf <- st_transform(
  assets_sf,
  st_crs(buildings)
)
```

## Visual Check

```{r}
# Get buildings dataset in form for leaflet map
buildings_leaflet <- buildings |>
  st_zm() |>
  st_cast("MULTIPOLYGON") 

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_sf,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```

## Join by FACLOCID number
```{r}
# Make polygons valid 
buildings_fixed <- st_make_valid(buildings_leaflet)

# Clean FACLOCID in buildings_fixed (to account for the duplicate id numbers that appear)
# From an examaninaton and cross checking the first number in the series is the current ID
buildings_fixed <- buildings_fixed |> 
  mutate(FACLOCID = sub(";.*", "", FACLOCID))

# Rename LocationID column to be FACLOCID for join
assets_facloc <- assets |>
  rename(FACLOCID = LocationID)

# Perform left join to match asset data to CONTRIBRES column by matching ID 
assets_facloc_joined <- assets_facloc |>
  left_join(buildings_fixed |>
              st_drop_geometry() |>
              # include RESNAME_CLEAN and BLDGALTNAME for JW check
              select(FACLOCID, CONTRIBRES, RESNAME_CLEAN, BLDGALTNAME), 
            by = "FACLOCID")

# View how many tagged
sum(!is.na(assets_facloc_joined$CONTRIBRES))
```

### Check FACLOCID join using a Jaro-Winkler Similarity test
In order to better confirm that the LocationID/FACLOCID match actually matched the correct assets together, the names of the buildings in the two datasets (Description_CLEAN and RESNAME_CLEAN) will be compared for a given row. The '_CLEAN' version of the building names will be used as it works better with the JW test

```{r}
# Filter assets_facloc_joined to rows where CONTRIBRES is not NA
assets_facloc_check <- assets_facloc_joined |>
  filter(!is.na(CONTRIBRES))

# Calculate Jaro-Winkler similarity between Description_CLEAN and RESNAME_CLEAN
assets_facloc_check <- assets_facloc_check |>
  mutate(jw_score = mapply(RecordLinkage::jarowinkler, Description_CLEAN, RESNAME_CLEAN))

# Filter to the Jaro-Winkler score of less than 0.90 for review
assets_facloc_review <- assets_facloc_check |>
  filter(jw_score < 0.90) |>
  select(
    FACLOCID,
    Description_CLEAN,
    RESNAME_CLEAN,
    
    jw_score,
    everything()
  )

View(assets_facloc_review)
```

FACLOCID numbers to flag: 
40834
Description: PARKING LOT BATTERY ALEXANDER
Resname: QUARTERS 804A

Found in data of this asset: "This feature is incorrectly described as "Battery Alexander parking lot" in FMSS."


## Spatial join for the untagged set

Note for self: This join will remove rows that do not contain lat lon points. I should be okay when I recombine everything at the end because these rows are included in the first ID join.
```{r}
# Clean buildings dataset to not include planned or proposed buildings 
buildings_fixed_current <- buildings_fixed |> 
  # Planned and proposed
   filter(!BLDGSTATUS %in% c("Proposed","Planned")) 
```


```{r}
# Filter to rows that do not contain contribres tag
assets_needing_spatial <- assets_facloc_joined |>
  filter(is.na(CONTRIBRES))


# Turn assets data into spatial points ----
# Filter out rows that are missing lat lon points 
assets_needing_spatial <- assets_needing_spatial|> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))

# Convert spreadsheet coordinates to sf points
assets_spatial_sf <- st_as_sf(
  assets_needing_spatial,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_spatial_sf <- st_transform(
  assets_sf,
  st_crs(buildings)
)

# Perform spatial join 
assets_spatial_join <- st_join(
  assets_spatial_sf,
  buildings_fixed_current |> 
    select(CONTRIBRES),
  join = st_within,
  left = TRUE
)

# Combine 2 CONTRIBRES columns 
assets_spatial_join <- assets_spatial_join |>
  mutate(CONTRIBRES = coalesce(CONTRIBRES.x, CONTRIBRES.y)) |> 
  # remove x and y column 
  select(c(-CONTRIBRES.x, -CONTRIBRES.y))

# View how many are still untagged (there should def be a decent amount from visual check in maps)
sum(!is.na(assets_spatial_join$CONTRIBRES))

```

