---
title: "asset-spatial-join"
format: html
---

```{r}
library(sf)
library(dplyr)
library(readxl)
library(leaflet)
library(writexl)
```


```{r}
# File paths

gdb_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/CCVA-asset-classifcation.gdb"

fc_name <- "GOGA_FACILITY_BuildingsAndStructures_py_20260520_ExportFeatures"

excel_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/goga-assets-cleaned.xlsx"

output_excel <- "goga-assets-cleaned-with-contributing-flag.xlsx"
```

```{r}
# MIGHT NOT USE
goga_buildings_path <- "C:/Users/jsibley/OneDrive - Golden Gate National Parks Conservancy/Documents/Code/projects/ccva/data/GOGA_Buildings.gdb"

GOGA_buildings <- st_read(
  dsn = goga_buildings_path,
  layer = "GOGA_FACILITY_Buildings_py",
  quiet = FALSE
)

# Check field names
names(GOGA_buildings)
```


```{r}
# Read polygon layer

buildings <- st_read(
  dsn = gdb_path,
  layer = fc_name,
  quiet = FALSE
)

# Transform CRS to WGS84 
buildings <- buildings |> 
  st_transform(4326) 

# Check field names
names(buildings)
```


```{r}
# Read asset spreadsheet
assets <- read_excel(excel_path, sheet = "Sheet1")

# Remove characters (+) from lat and lon columns and convert to numeric
assets <- assets |> 
  mutate(
    PrimaryLatitude = gsub("\\+", "", PrimaryLatitude),
    PrimaryLongitude = gsub("\\+", "", PrimaryLongitude))|>
  mutate(PrimaryLatitude = as.numeric(PrimaryLatitude),
         PrimaryLongitude = as.numeric(PrimaryLongitude)) 

# Filter rows without lat lon points 
assets_point <- assets |> 
  filter(!is.na(PrimaryLatitude) | !is.na(PrimaryLongitude))
```


```{r}
# Identify rows with missing lat lon points 
assets_missing_point <- assets |> 
  filter(is.na(PrimaryLatitude) | is.na(PrimaryLongitude))

# See if missing rows are in GOGA_Building via Location ID
matches_GOGA_build <- assets_missing_point |> 
  filter(Location %in% GOGA_buildings$FACLOCID)

# See if missing rows are in buildings dataset via Location ID
matches_buildings <- assets_missing_point |> 
  filter(Location %in% buildings$FACLOCID)
```

80 rows within assets dataset do not have a lat lon point. Of those 80, only 7 have a matching Location ID within GOGA_Buildings dataset and 1 has a matching Location ID within buildings dataset.  Need to decide if its worth it? Might be wiser to go through and assign lat lon point to missing rows 

```{r}
# Convert spreadsheet coordinates to sf points
assets_sf <- st_as_sf(
  assets_point,
  coords = c("PrimaryLongitude", "PrimaryLatitude"),
  crs = 4326,      # WGS84 lat/long
  remove = FALSE
)

# Reproject points to polygon CRS (WGS84)
assets_sf <- st_transform(
  assets_sf,
  st_crs(buildings)
)
```

```{r}
# Get buildings dataset in form for leaflet map
buildings_leaflet <- buildings |>
  st_zm() |>
  st_cast("MULTIPOLYGON") 

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_sf,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```


```{r}
# Spatial join

# First make valid 
buildings_fixed <- st_make_valid(buildings_leaflet)

# Perform spatial join 
assets_joined <- st_join(
  assets_sf,
  buildings_fixed |> 
    select(CONTRIBRES),
  join = st_within,
  left = TRUE
)

# Determine how many rows are still missing contributing resource flag
assets_joined_missing <- assets_joined |> 
  filter(is.na(CONTRIBRES))
nrow(assets_joined_missing)
```
681 rows out of 1332 (51%) did not receive the contributing resource tag from the spatial join.


```{r}
# Recreate map with only missed tagged points 

# Create map
#--------------------------------------------------
leaflet() |> 
  
  addProviderTiles(providers$Esri.WorldImagery, group = "Aerial") |> 
  
  addProviderTiles(providers$CartoDB.Positron, group = "Basemap") |> 
  
  addPolygons(
    data = buildings_leaflet,
    color = "blue",
    weight = 1,
    fillOpacity = 0.3,
     popup = ~ paste(
       "<b>Resource Name:</b>", RESNAME,
       "<b>Contributing Resource Flag:</b>", CONTRIBRES
    ),
    group = "Resource Polygons"
  ) |> 
  
   addCircleMarkers(
    data = assets_joined_missing,
    radius = 4,
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~ paste(
      "<b>Description:</b>", Description,
      "<b>Latitude:</b>",PrimaryLatitude,
      "<br><b>Longitude:</b>",PrimaryLongitude
    ),
    group = "Asset Points"
   ) |> 
  
   addLayersControl(
     overlayGroups = c("Resource Polygons", "Asset Points"),
     baseGroups = c("Aerial", "Basemap"),
     options = layersControlOptions(collapsed = FALSE)
  )

```

Next step: 

Try one more join (of just the missing points) with FACLOCID to see if there are a few more that could get labeled. 

HAVE NOT TRIED PAST THIS
```{r}
# Remove geometry before exporting
#assets_output <- assets_joined %>%
#  st_drop_geometry()
```


```{r}
# Write result to Excel
#write_xlsx(
#  assets_output,
#  output_excel
#)

#cat("Done. Output saved to:", output_excel)
```

